\( \tool \): JavaScript IR-based Semantics Extraction Toolchain

### 2020-04-02 ase20/abstract.tex
Forward compatibility by Stage 4 proposals not by ES11

To alleviate the problem, we propose \( \tool \), a JavaScript IR-based
Semantics Extraction Toolchain.  It is the first tool that \textit{automatically
synthesizes} parsers and AST-IR translators directly from a given
language specification, ECMAScript.
For syntax, we develop a parser generation technique with \textit{lookahead
parsing} for \( \bnfes \), a variant of the extended BNF used in ECMAScript.
For semantics, \( \tool \) synthesizes AST-IR translators using \textit{forward
compatible} rule-based compilation.  \textit{Compile rules} describe how to
convert each step of abstract algorithms written in a structured natural language
into \( \ires \), an Intermediate Representation that we designed for
ECMAScript.  For the five most recent ECMAScript versions, \( \tool \) automatically
synthesized parsers for all versions, and compiled \inred{XX.XX}\% of the
algorithm steps on average.  After we complete the missing parts manually,
the extracted core semantics of the latest ECMAScript version (ES10, 2019)
passed all \inred{XX,XXX} applicable tests.
Using this \textit{first formal semantics of modern JavaScript}, we found
\inred{X} specification errors in ES10, which were all confirmed by
the Ecma Technical Committee 39.  Furthermore, we also successfully
applied \( \tool \) to the current draft of the next ECMAScript (ES11, 2020),
which shows the forward compatibility of \( \tool \).
\end{abstract}

>>>>

To alleviate the problem, we propose \( \tool \), a JavaScript IR-based
Semantics Extraction Toolchain.  It is the first tool that \textit{automatically
synthesizes} parsers and AST-IR translators directly from a given
language specification, ECMAScript.
For syntax, we develop a parser generation technique with \textit{lookahead
parsing} for \( \bnfes \), a variant of the extended BNF used in ECMAScript.
For semantics, \( \tool \) synthesizes AST-IR translators using \textit{forward
compatible} rule-based compilation.  \textit{Compile rules} describe how to
convert each step of abstract algorithms written in a structured natural language
into \( \ires \), an Intermediate Representation that we designed for
ECMAScript.  For the five most recent ECMAScript versions, \( \tool \) automatically
synthesized parsers for all versions, and compiled \inred{XX.XX}\% of the
algorithm steps on average.  After we complete the missing parts manually,
the extracted core semantics of the latest ECMAScript version (ES10, 2019)
passed all \inred{XX,XXX} applicable tests.
Using this \textit{first formal semantics of modern JavaScript}, we found
\inred{X} specification errors in ES10, which were all confirmed by
the Ecma Technical Committee 39.  Furthermore, we also successfully
applied \( \tool \) to \inred{X} feature proposals that are ready
for inclusion in the next ECMAScript, 
which shows the forward compatibility of \( \tool \).
\end{abstract}

### 2020-04-02 ase20/intro.tex
Forward compatibility by Stage 4 proposals not by ES11

Despite its popularity, JavaScript developers often suffer from its intricate
semantics, which may cause unexpected behaviors.  For example, calling the
following JavaScript function seems to always return \( \code{false} \):
\begin{lstlisting}[style=myJSstyle]
        function f(x) { return x == !x; }
\end{lstlisting}
Unfortunately, it returns \( \code{true} \) when its argument
is an empty array \( \code{[]} \).  To correctly understand and reason about
such a complex behavior, the formal semantics of JavaScript is necessary.

@@

However, to the best of our knowledge, all existing approaches to JavaScript
IR-based semantics extraction \textit{manually} build parsers and translators.
Although manually building them was reasonable until ECMAScript 5.1 (ES5.1,
2011)~\cite{es5}, it is too tedious, labor-intensive, and
error-prone to deal with the large size of \textit{modern JavaScript} since ECMAScript 6
(ES6, 2015)~\cite{es6}. ES6 introduced numerous significantly new features such
as lexical binding via \( \code{let} \), the spread \( \code{...} \) operator,
classes, the \( \code{for-of} \) operator, the \( \code{async} \) functions, and
generators.  For example, consider KJS~\cite{kjs}.
KJS is one of formal semantics of ES5.1
and it is defined on top of \( \kframework \), which is a framework for
defining language semantics.  According to an author of KJS, it took
\textit{four months} to implement an AST-IR translator for 1,370 steps out of 2,932
steps in 368 abstract algorithms~\cite{kjsslides}. However, the most recent version of ECMAScript
(ES10, 2019)~\cite{es10} has 1,731 abstract algorithms consisting of 10,068
steps. Thus, the manual approaches do not seem to be scalable enough to build
an AST-IR translator for modern JavaScript,
and indeed no formal semantics exists for ES6 to ES10.

@@

To alleviate this problem, we propose a technique to \textit{automatically
synthesize} parsers and AST-IR translators directly from ECMAScript with
\textit{forward compatibility}. There are several technical challenges in
synthesizing parsers and translators. For syntax, ECMAScript utilizes its own
special variant of EBNF with parametric non-terminals, conditional
alternatives, and various special terminal symbols.  Thus, no existing parser
generation technique is directly applicable for this variant.  Moreover,
JavaScript provides automatic semicolon insertion in its parsing algorithm with
several complex rules, not in a lexer. For semantics, abstract algorithms in
ECMAScript are written in English.  Besides, a general and forward compatible
representation of abstract algorithms is necessary to support future
versions of ECMAScript.

Our contribution is \( \tool \), a JavaScript IR-based Semantics Extraction
Toolchain:
\begin{itemize}[leftmargin=0.5cm]
  \item \textbf{\( \tool \) is the \textit{first tool that can automatically synthesize}
    parsers and AST-IR translators directly from a language specification,
    ECMAScript.}  For syntax, we formally introduce a variant of EBNF, \( \bnfes \),
    and propose a parser generation technique with
    lookahead parsing for \( \bnfes \), which supports automatic semicolon
    insertion. For semantics, we propose semi-automatic synthesis of AST-IR
    translators assisted by compile rules because of the
    difficulty in handling the semantics written in a natural language.  Compile rules describe how
    each step of abstract algorithms is converted into our intermediate
    representation \( \ires \) designed for ECMAScript. We evaluated \( \tool \)
    with the five most recent ECMAScript versions (ES6 to ES10).   \( \tool \) automatically generated
    parsers for all versions, and automatically compiled
    \inred{XX.X}\% of the steps in abstract algorithms on average.
  \item \textbf{ Using \( \tool \), we could define the \textit{first IR-based
    formal semantics of modern JavaScript}.}  The semantics targets the
    core of ES10, and \( \tool \) successfully synthesizes its
    parser and AST-IR translator for \inred{X,XXX} out of \inred{X,XXX}
    algorithm steps.  We manually completed the missing parts of the AST-IR
    translator (\inred{XXX} steps) in \inred{XX} days.  We evaluated the correctness
    of the formal semantics using Test262~\cite{test262}, the official
    conformance test suite, and the semantics passed all \inred{XX,XXX}
    applicable tests.  Moreover, we found \inred{X} specification errors in ES10
    thanks to the formal semantics.  They were all confirmed by TC39 and will be
    fixed in the next release.
  \item \textbf{\( \tool \) is also \textit{forward compatible} with new language features
    proposed for future ECMAScript specifications.}  
    We evaluated the forward compatibility of \( \tool \) by applying it
    to the current draft of the next ECMAScript (ES11, 2020).  It successfully
    synthesized a parser and automatically
    compiled \inred{XX} out of \inred{XX} algorithm steps for ES11.
    After completion of the missing parts, the extracted semantics passed all \inred{XX} tests.
\end{itemize}

>>>>

Despite its popularity, JavaScript developers often suffer from its intricate
semantics, which may cause unexpected behaviors.  For example, the
following function may seem to always return \( \code{false} \):
\begin{lstlisting}[style=myJSstyle]
        function f(x) { return x == !x; }
\end{lstlisting}
Unfortunately, it returns \( \code{true} \) when its argument
is an empty array \( \code{[]} \).  To correctly understand and reason about
such a complex behavior, the formal semantics of JavaScript is necessary.

@@

However, to the best of our knowledge, all existing approaches to JavaScript
IR-based semantics extraction \textit{manually} build parsers and translators.
Although manually building them was reasonable until ECMAScript 5.1 (ES5.1,
2011)~\cite{es5}, it is too tedious, labor-intensive, and
error-prone to deal with the large size of \textit{modern JavaScript} since ECMAScript 6
(ES6, 2015)~\cite{es6}. ES6 introduced numerous significantly new features such
as lexical binding via \( \code{let} \), the spread \( \code{...} \) operator,
classes, the \( \code{for-of} \) operator, the \( \code{async} \) functions, and
generators.  For example, consider KJS~\cite{kjs},
one of formal semantics of ES5.1 defined on top of \( \kframework \), which is
a framework for defining language semantics.
According to an author of KJS, it took
\textit{four months} to implement an AST-IR translator for 1,370 steps out of 2,932
steps in 368 abstract algorithms~\cite{kjsslides}. However, the most recent version of ECMAScript
(ES10, 2019)~\cite{es10} has 1,731 abstract algorithms consisting of 10,068
steps. Thus, the manual approaches do not seem to be scalable enough to build
an AST-IR translator for modern JavaScript,
and indeed no formal semantics exists for ES6 to ES10.

@@

To alleviate this problem, we propose a technique to \textit{automatically
synthesize} parsers and AST-IR translators directly from ECMAScript with
\textit{forward compatibility}. There are several technical challenges in
synthesizing parsers and translators. For syntax, ECMAScript utilizes its own
variant of EBNF with parametric non-terminals, conditional
alternatives, and various special terminal symbols.  Thus, no existing parser
generation technique is directly applicable for this variant.  Moreover,
JavaScript provides automatic semicolon insertion in its parsing algorithm with
several complex rules, not in a lexer. For semantics, abstract algorithms in
ECMAScript are written in English.  Besides, a general and forward compatible
representation of abstract algorithms is necessary to support future
versions of ECMAScript.

Our contribution is \( \tool \), a JavaScript IR-based Semantics Extraction
Toolchain:
\begin{itemize}[leftmargin=0.5cm]
  \item \textbf{\( \tool \) is the \textit{first tool that can automatically synthesize}
    parsers and AST-IR translators directly from a language specification,
    ECMAScript.}  For syntax, we formally introduce a variant of EBNF, \( \bnfes \),
    and propose a parser generation technique with
    lookahead parsing for \( \bnfes \), which supports automatic semicolon
    insertion. For semantics, we propose semi-automatic synthesis of AST-IR
    translators assisted by compile rules due to the
    difficulty in handling the semantics written in a natural language.  Compile rules describe
    how to convert each step of abstract algorithms into our intermediate
    representation \( \ires \) designed for ECMAScript. We evaluated \( \tool \)
    with the five most recent ECMAScript versions (ES6 to ES10).   \( \tool \) automatically generated
    parsers for all versions, and automatically compiled
    \inred{XX.X}\% of the steps in abstract algorithms on average.
  \item \textbf{ Using \( \tool \), we could define the \textit{first IR-based
    formal semantics of modern JavaScript}.}  The semantics targets the
    core of ES10, and \( \tool \) successfully synthesizes its
    parser and AST-IR translator for \inred{X,XXX} out of \inred{X,XXX}
    algorithm steps.  We manually completed the missing parts of the AST-IR
    translator (\inred{XXX} steps) in \inred{XX} days.  We evaluated the correctness
    of the formal semantics using Test262~\cite{test262}, the official
    conformance test suite, and the semantics passed all \inred{XX,XXX}
    applicable tests.  Moreover, we found \inred{X} specification errors in ES10
    thanks to the formal semantics.  They were all confirmed by TC39 and will be
    fixed in the next release.
  \item \textbf{\( \tool \) is also \textit{forward compatible} with new language features
    proposed for future ECMAScript specifications.}  
     We evaluated the forward compatibility of \( \tool \) by applying it
     to all \inred{XX} proposals that are ready for inclusion in the next ECMAScript  (ES11, 2020).
    It automatically synthesized parsers and
    compiled \inred{XX} out of \inred{XX} algorithm steps for all the proposals.
    After completion of the missing parts, the extracted semantics passed all \inred{XX} tests.
\end{itemize}
